Finding shortest and closest vectors in a lattice of Voronoi's first kind

### 2014-04-21 paper.tex
minor

Equivalently the Voronoi cell can be defined as the intersection of the half spaces 
\begin{align*}
H_{v} &= \{x \in \reals^n \mid \|x\| \leq \|x - v\| \} \\
&= \{x \in \reals^n \mid \dotprod{x}{v} \leq \tfrac{1}{2}\dotprod{v}{v} \}
\end{align*}
for all $v \in \Lambda \backslash  \{\zerobf\}$.  %We denote by $x^\prime$ the transpose of the vector $x$ and so $\dotprod{x}{v}$ is the inner product between column vectors $x$ and $v$.  
We denote by $\dotprod{x}{v}$ the inner product between vectors $x$ and $v$.
It is not necessary to consider all $v \in \Lambda \backslash  \{\zerobf\}$ to define the Voronoi cell.   %Figure \ref{lattices:fig:vorregion} depicts the relevant vectors of the lattice with generator given by~\eqref{lattices:eq:genB}.  A lattice point in $\calR$ is called \term{relevant}. The following remark follow trivially from the preceding discussion.
The \emph{relevant vectors} $v \in \Lambda$ are those for which  
\[
\dotprod{v}{x} \leq \dotprod{x}{x}
\]
for all $x \in \Lambda$ with $x \neq v$ and $x \neq \zerobf$.  We denote by $\relevant(\Lambda)$ the set of relevant vectors of the lattice $\Lambda$.  The Voronoi cell is the intersection of the halfspaces corresponding with the relevant vectors, that is, 
\[
%\begin{equation}\label{eq:defvortermsrelvecs}
\vor(\Lambda) = \cap_{v\in\relevant(\Lambda)}{H_{v}}.
%\end{equation}
\]
The closest lattice point problem and the Voronoi cell are related in that $x\in\Lambda$ is a closest lattice point to $y$ if and only $y - x \in \vor(\Lambda)$, that is, if and only if
\begin{equation}\label{eq:relvectnearpointieq}
\dotprod{(y - x)}{v} \leq \tfrac{1}{2} \dotprod{v}{v}
\end{equation} 
for all $v \in \relevant(\Lambda)$.

>>>>

Equivalently the Voronoi cell can be defined as the intersection of the half spaces 
\begin{align*}
H_{v} &= \{x \in \reals^n \mid \|x\| \leq \|x - v\| \} \\
&= \{x \in \reals^n \mid \dotprod{x}{v} \leq \tfrac{1}{2}\dotprod{v}{v} \}
\end{align*}
for all $v \in \Lambda \backslash  \{\zerobf\}$.  %We denote by $x^\prime$ the transpose of the vector $x$ and so $\dotprod{x}{v}$ is the inner product between column vectors $x$ and $v$.  
We denote by $\dotprod{x}{v}$ the inner product between vectors $x$ and $v$.
It is not necessary to consider all $v \in \Lambda \backslash  \{\zerobf\}$ to define the Voronoi cell.   %Figure \ref{lattices:fig:vorregion} depicts the relevant vectors of the lattice with generator given by~\eqref{lattices:eq:genB}.  A lattice point in $\calR$ is called \term{relevant}. The following remark follow trivially from the preceding discussion.
The \emph{relevant vectors} are those lattice points $v \in \Lambda \backslash  \{\zerobf\}$ for which  
\[
\dotprod{v}{x} \leq \dotprod{x}{x} \qquad \text{for all $x \in \Lambda$}.
\]
We denote by $\relevant(\Lambda)$ the set of relevant vectors of the lattice $\Lambda$.  The Voronoi cell is the intersection of the halfspaces corresponding with the relevant vectors, that is, 
\[
%\begin{equation}\label{eq:defvortermsrelvecs}
\vor(\Lambda) = \cap_{v\in\relevant(\Lambda)}{H_{v}}.
%\end{equation}
\]
The closest lattice point problem and the Voronoi cell are related in that $x\in\Lambda$ is a closest lattice point to $y$ if and only $y - x \in \vor(\Lambda)$, that is, if and only if
\begin{equation}\label{eq:relvectnearpointieq}
\dotprod{(y - x)}{v} \leq \tfrac{1}{2} \dotprod{v}{v} \qquad \text{for all $v \in \relevant(\Lambda)$}.  
\end{equation}

### 2014-04-21 talk/presentation.tex
minor

The \term{Voronoi cell} of a lattice $\Lambda \subset \reals^m$ is the subset of $\reals^m$ with points at least as close to the origin than to any lattice point,
\[
\vor(\Lambda) = \{ x \in \reals^m \mid \|x\| \leq \|x - y\|, y \in \Lambda \}.
\]

@@

\frame{ 
\frametitle{Relevant vectors}
\begin{itemize}
\item The set of relevant vectors $\relevant(\Lambda)$ contains those $v \in \Lambda$ such that  
\[
\dotprod{v}{x} \leq \dotprod{x}{x}
\]
for all $x \in \Lambda$ with $x \neq v$ and $x \neq \zerobf$.  
\item The Voronoi cell can be defined using the relevant vectors, 
\[
\vor(\Lambda) = \{ x \in\reals^m \mid \|x\| \leq \|x - v\|, v \in\relevant(\Lambda) \}
\]
\item Short vectors are relevant vectors.
\end{itemize}
}

@@

Given a lattice $\Lambda \subset \reals^m$ and a vector $y \in \reals^m$, a problem of interest is to find $x \in \Lambda$ such that
\[
\| y - x \|^2
\] 
is minimised.

@@

Let $x_0$ be some lattice point from $\Lambda$ and consider the following iteration,
\begin{align*}
x_{k+1} &= x_k + v_k \\
v_k &= \arg\min_{ v \in \relevant(\Lambda) \cup \{\zerobf\} } \|y - x_k - v \|
\end{align*} 
}

@@

\frame{
\frametitle{A series of relevant vectors}
\begin{itemize}
\item The number $K$ of iterations might be large.
\item The initial point $x_0$ affects $K$.
\item Minimising over the set of relevant vectors, that is computing
\[
\arg\min_{ v \in \relevant(\Lambda) \cup \{\zerobf\} } \|y - x_k - v \|
\]
might be expensive.
\item There are as many as $2^{n+1}-2$ relevant vectors.
\end{itemize}
}

>>>>

The \term{Voronoi cell} of a lattice $\Lambda \subset \reals^m$ is the subset of $\reals^m$ at least as close to the origin than to any lattice point,
\[
\vor(\Lambda) = \{ x \in \reals^m \mid \|x\| \leq \|x - y\|, y \in \Lambda \}.
\]

@@

\frame{ 
\frametitle{Relevant vectors}
\begin{itemize}
\item The relevant vectors are those $v \in \Lambda \backslash \{0\}$ such that  
\[
\dotprod{v}{x} \leq \dotprod{x}{x} \qquad \text{for all $x \in \Lambda$.}
\]
\item The Voronoi cell can be defined using the relevant vectors, 
\[
\vor(\Lambda) = \{ x \in\reals^m \mid \|x\| \leq \|x - v\|, v \in\relevant(\Lambda) \}.
\]
\item Short vectors are relevant vectors.
\end{itemize}
}

@@

Given a lattice $\Lambda \subset \reals^m$ and a vector $y \in \reals^m$ find $x \in \Lambda$ such that
\[
\| y - x \|^2
\] 
is minimised.

@@

Let $x_0$ be some lattice point from $\Lambda$ and consider the following iteration,
\begin{align*}
x_{k+1} &= x_k + v_k \\
v_k &= \arg\min_{ v \in \relevant(\Lambda) \cup \{0\} } \|y - x_k - v \|
\end{align*} 
}

@@

\frame{
\frametitle{A series of relevant vectors}
\begin{itemize}
\item The number $K$ of iterations might be large.
\item The initial point $x_0$ affects $K$.
\item Minimising over the set of relevant vectors, that is computing
\[
\arg\min_{ v \in \relevant(\Lambda) \cup \{0\} } \|y - x_k - v \|
\]
might be expensive.
\item There are as many as $2^{n+1}-2$ relevant vectors.
\end{itemize}
}
